---
title: "Practical Machine Learning Week 4 Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(reshape2))
suppressMessages(library(caret))
suppressMessages(library(gridExtra))
suppressMessages(library(rattle))
```

```{r, echo=FALSE}
set.seed(123)
training<-read.csv("pml-training.csv")
testing<-read.csv("pml-testing.csv")

training[is.na(training)]<-0
cormatrix<-cor(training[,c(18,21,37,38,39,40:48,64:66)])

```

```{r, echo=FALSE}
##Near zero variation variables
nsv<-nearZeroVar(training,saveMetrics = TRUE)
```

```{r, echo=FALSE}
AccelBelt<-training %>% group_by(classe) %>% summarise(average = mean(total_accel_belt))
AccelArm<-training %>% group_by(classe) %>% summarise(average = mean(total_accel_arm))
AccelDB<-training %>% group_by(classe) %>% summarise(average = mean(total_accel_dumbbell))
AccelForearm<-training %>% group_by(classe) %>% summarise(average = mean(total_accel_forearm))

B<-ggplot(AccelBelt,aes(x=classe,y=average,colour=classe))+geom_point(size=5)+theme_minimal()
A<-ggplot(AccelArm,aes(x=classe,y=average,colour=classe))+geom_point(size=5)+theme_minimal()
D<-ggplot(AccelDB,aes(x=classe,y=average,colour=classe))+geom_point(size=5)+theme_minimal()
F<-ggplot(AccelForearm,aes(x=classe,y=average,colour=classe))+geom_point(size=5)+theme_minimal()

#grid.arrange(B,A,D,F, ncol=2)

```

#Look at boxplots (Classe A has much lower median for total belt acceleration)
```{r, echo=FALSE}
AccelBeltBP<-ggplot(training,aes(x=classe,y=total_accel_belt,color=classe))+geom_boxplot()
AccelArmBP<-ggplot(training,aes(x=classe,y=total_accel_arm,color=classe))+geom_boxplot()
AccelDBBP<-ggplot(training,aes(x=classe,y=total_accel_dumbbell,color=classe))+geom_boxplot()
AccelForearmBP<-ggplot(training,aes(x=classe,y=total_accel_forearm,color=classe))+geom_boxplot()

grid.arrange(AccelBeltBP,AccelArmBP,AccelDBBP,AccelForearmBP, ncol=2)
```

#Look at boxplots for magnet_x (Classe A has much lower median for magnet arm. Classe E high variability for belt)
```{r, echo=FALSE}
MagnetXBelt<-ggplot(training,aes(x=classe,y=magnet_belt_x,color=classe))+geom_boxplot()
MagnetXArm<-ggplot(training,aes(x=classe,y=magnet_arm_x,color=classe))+geom_boxplot()
MagnetXDB<-ggplot(training,aes(x=classe,y=magnet_dumbbell_x,color=classe))+geom_boxplot()
MagnetXForearm<-ggplot(training,aes(x=classe,y=magnet_forearm_x,color=classe))+geom_boxplot()

grid.arrange(MagnetXBelt,MagnetXArm,MagnetXDB,MagnetXForearm)
```

```{r, echo=FALSE}
MagnetYBelt<-ggplot(training,aes(x=classe,y=magnet_belt_y,color=classe))+geom_boxplot()
MagnetYArm<-ggplot(training,aes(x=classe,y=magnet_arm_y,color=classe))+geom_boxplot()
MagnetYDB<-ggplot(training,aes(x=classe,y=magnet_dumbbell_y,color=classe))+geom_boxplot()
MagnetYForearm<-ggplot(training,aes(x=classe,y=magnet_forearm_y,color=classe))+geom_boxplot()

#grid.arrange(MagnetYBelt,MagnetYArm,MagnetYDB,MagnetYForearm)
```

#Look at boxplots for magnet_z (Classe E high variability for belt)
```{r, echo=FALSE}
MagnetZBelt<-ggplot(training,aes(x=classe,y=magnet_belt_z,color=classe))+geom_boxplot()
MagnetZArm<-ggplot(training,aes(x=classe,y=magnet_arm_z,color=classe))+geom_boxplot()
MagnetZDB<-ggplot(training,aes(x=classe,y=magnet_dumbbell_z,color=classe))+geom_boxplot()
MagnetZForearm<-ggplot(training,aes(x=classe,y=magnet_forearm_z,color=classe))+geom_boxplot()

grid.arrange(MagnetZBelt,MagnetZArm,MagnetZDB,MagnetZForearm)
```

#Look at gyros x (Classe E high variability for belt. Classe A high variability for arm)
```{r, echo=FALSE}
GyrosXBelt<-ggplot(training,aes(x=classe,y=gyros_belt_x,color=classe))+geom_boxplot()
GyrosXArm<-ggplot(training,aes(x=classe,y=gyros_arm_x,color=classe))+geom_boxplot()
GyrosXDB<-ggplot(training,aes(x=classe,y=gyros_dumbbell_x,color=classe))+geom_boxplot()
GyrosXForearm<-ggplot(training,aes(x=classe,y=gyros_forearm_x,color=classe))+geom_boxplot()

grid.arrange(GyrosXBelt,GyrosXArm,GyrosXDB,GyrosXForearm)
```

```{r, echo=FALSE}
GyrosYBelt<-ggplot(training,aes(x=classe,y=gyros_belt_y,color=classe))+geom_boxplot()
GyrosYArm<-ggplot(training,aes(x=classe,y=gyros_arm_y,color=classe))+geom_boxplot()
GyrosYDB<-ggplot(training,aes(x=classe,y=gyros_dumbbell_y,color=classe))+geom_boxplot()
GyrosYForearm<-ggplot(training,aes(x=classe,y=gyros_forearm_y,color=classe))+geom_boxplot()

#grid.arrange(GyrosYBelt,GyrosYArm,GyrosYDB,GyrosYForearm)
```

```{r, echo=FALSE}
GyrosZBelt<-ggplot(training,aes(x=classe,y=gyros_belt_z,color=classe))+geom_boxplot()
GyrosZArm<-ggplot(training,aes(x=classe,y=gyros_arm_z,color=classe))+geom_boxplot()
GyrosZDB<-ggplot(training,aes(x=classe,y=gyros_dumbbell_z,color=classe))+geom_boxplot()
GyrosZForearm<-ggplot(training,aes(x=classe,y=gyros_forearm_z,color=classe))+geom_boxplot()

#grid.arrange(GyrosZBelt,GyrosZArm,GyrosZDB,GyrosZForearm)
```

#Look at roll (Classe A much lower median for Belt. Classe C much higher median for forearm)
```{r, echo=FALSE}
RollBelt<-ggplot(training,aes(x=classe,y=roll_belt,colour=classe))+geom_boxplot()
RollArm<-ggplot(training,aes(x=classe,y=roll_arm,colour=classe))+geom_boxplot()
RollDumbbell<-ggplot(training,aes(x=classe,y=roll_dumbbell,colour=classe))+geom_boxplot()
RollForearm<-ggplot(training,aes(x=classe,y=roll_forearm,colour=classe))+geom_boxplot()

grid.arrange(RollBelt,RollArm,RollDumbbell,RollForearm)
```

```{r, echo=FALSE}
PitchBelt<-ggplot(training,aes(x=classe,y=pitch_belt,colour=classe))+geom_boxplot()
PitchArm<-ggplot(training,aes(x=classe,y=pitch_arm,colour=classe))+geom_boxplot()
PitchDumbbell<-ggplot(training,aes(x=classe,y=pitch_dumbbell,colour=classe))+geom_boxplot()
PitchForearm<-ggplot(training,aes(x=classe,y=pitch_forearm,colour=classe))+geom_boxplot()

#grid.arrange(PitchBelt,PitchArm,PitchDumbbell,PitchForearm)
```

#Look at yaw (Classe A much lower for belt. Class C much higher for Forearm)
```{r, echo=FALSE}
YawBelt<-ggplot(training,aes(x=classe,y=yaw_belt,colour=classe))+geom_boxplot()
YawArm<-ggplot(training,aes(x=classe,y=yaw_arm,colour=classe))+geom_boxplot()
YawDumbbell<-ggplot(training,aes(x=classe,y=yaw_dumbbell,colour=classe))+geom_boxplot()
YawForearm<-ggplot(training,aes(x=classe,y=yaw_forearm,colour=classe))+geom_boxplot()

grid.arrange(YawBelt,YawArm,YawDumbbell,YawForearm)
```

```{r, echo=FALSE}
ARB<-training %>% group_by(classe) %>% summarise(average = mean(amplitude_roll_belt))
ARA<-training %>% group_by(classe) %>% summarise(average = mean(amplitude_roll_arm))
ARD<-training %>% group_by(classe) %>% summarise(average = mean(amplitude_roll_dumbbell))
ARF<-training %>% group_by(classe) %>% summarise(average = mean(amplitude_roll_forearm))

AmpRollBelt<-ggplot(ARB,aes(x=classe,y=average,colour=classe))+geom_point(size = 5)+labs(y="Belt")
AmpRollArm<-ggplot(ARA,aes(x=classe,y=average,colour=classe))+geom_point(size = 5)+labs(y="Arm")
AmpRollDumbbell<-ggplot(ARD,aes(x=classe,y=average,colour=classe))+geom_point(size = 5)+labs(y="Dumbbell")
AmpRollForearm<-ggplot(ARF,aes(x=classe,y=average,colour=classe))+geom_point(size = 5)+labs(y="Forearm")

#grid.arrange(AmpRollBelt,AmpRollArm,AmpRollDumbbell,AmpRollForearm)
```

```{r, echo=FALSE}
APB<-training %>% group_by(classe) %>% summarise(average = mean(amplitude_pitch_belt))
APA<-training %>% group_by(classe) %>% summarise(average = mean(amplitude_pitch_arm))
APD<-training %>% group_by(classe) %>% summarise(average = mean(amplitude_pitch_dumbbell))
APF<-training %>% group_by(classe) %>% summarise(average = mean(amplitude_pitch_forearm))

AmpPitchBelt<-ggplot(APB,aes(x=classe,y=average,colour=classe))+geom_point(size = 5)+labs(y="Belt")
AmpPitchArm<-ggplot(APA,aes(x=classe,y=average,colour=classe))+geom_point(size = 5)+labs(y="Arm")
AmpPitchDumbbell<-ggplot(APD,aes(x=classe,y=average,colour=classe))+geom_point(size = 5)+labs(y="Dumbbell")
AmpPitchForearm<-ggplot(APF,aes(x=classe,y=average,colour=classe))+geom_point(size = 5)+labs(y="Forearm")

#grid.arrange(AmpPitchBelt,AmpPitchArm,AmpPitchDumbbell,AmpPitchForearm)
```
#First model
```{r}
model<-train(classe~roll_belt+yaw_belt+total_accel_belt+amplitude_roll_belt+amplitude_pitch_belt+gyros_belt_x+gyros_belt_z+amplitude_roll_belt+amplitude_pitch_belt+gyros_belt_x+gyros_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+gyros_arm_x+magnet_arm_x+amplitude_roll_arm+amplitude_pitch_arm+amplitude_roll_dumbbell+amplitude_pitch_dumbbell+roll_forearm+yaw_forearm+amplitude_roll_forearm+amplitude_pitch_forearm,data=training,method="rf")

varImpPlot(model$finalModel)
```

#Second model with cross validation, taking out irrelevant variables
```{r}
train_control<-trainControl(method="cv", number=10, savePredictions=TRUE)
model2<-train(classe~roll_belt+yaw_belt+total_accel_belt+gyros_belt_x+gyros_belt_z++gyros_belt_x+gyros_belt_z+magnet_belt_x+magnet_belt_y+magnet_belt_z+gyros_arm_x+magnet_arm_x+roll_forearm+yaw_forearm,data=training,method="rf",trControl = train_control)

model2$finalModel
```

#Prediction
```{r}
pred<-predict(model2,newdata = testing)
pred
```

